Fisheries management when quotas are costly to change

### 2014-07-24 manuscripts/presentation.Rmd
skeleton for esa slides

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviera
    affiliation: UTK
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTK

@@

abstract: |
          Ecological systems are dynamic and policies to manage them need to
          respond to that variation. However, policy adjustments will sometimes
          be costly, which means that fine-tuning a policy to track variability in
          the environment very tightly will only sometimes be worthwhile. We use a
          classic fisheries management question -- how to manage a stochastically
          varying population using annually varying quotas in order to maximize
          profit  - to examine how costs of policy adjustment change optimal
          management recommendations. Costs of policy adjustment (here changes in
          fishing quotas through time) could take different forms. For example,
          these costs may respond to the size of the change being implemented or
          there could be a fixed cost any time a quota change is made. We show
          how different forms of policy costs have contrasting implications for
          optimal policies. While some types of cost act to smooth out variation in
          quotas and stock sizes, as one might have expected, others can actually
          increase variation in stock sizes and quotas through time. We also show
          that the potential economic impact on the fishery of managers assuming
          policy adjustment costs are present when in fact they are absent is much
          smaller than the impact of managers assuming such costs are absent if
          they are present.

layout: review,12pt
linenumbers: true

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

output:
  pdf_document:
    template: components/elsarticle.latex
    fig_caption: true
---

@@

library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

@@

Introduction
============

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2010]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to reflect environmental variation more closely
should be traded off against potential costs associated with the more
interventionist approach to management this would require.

```{r Figure_1, fig.cap="Quota on Blue Fin Tuna Catch", fig.width=4, fig.height=4}
tuna <- read.csv("components/data/tuna.csv")
ggplot(tuna, aes(year, quota)) + geom_point()
```

To illustrate these concepts, Figure 1 show an example from fisheries
management. On the left axis, the figure shows a time series for the
estimated population size, represented as spawning stock biomass, of
west Atlantic bluefin tuna (_Thunnus thynnus_, henceforth bluefin) from
a recent stock assessment [@ICCAT20XX].  On the right axis the figure
also shows the catch quota for the stock that was set by the relevant
management agency [@ICCAT20XX]. Despite the estimated population size
declining by XX% between YYYY and ZZZZ, the quota was not changed during
this period. Instead, the quota only changed occasionally and in between
times was left unaltered. Fishery management decisions regarding this
species can be highly contentious (@Safina1998, @Sissenwine1998,
@Porch2005). Moreover, the stock is fished by fleets from many nations
with quotas being set by a multilaterial management agency through a
process of negotiation. As such, we might reasonably anticipate that
for this species there could be substantial transaction costs involved
in reaching agreement over any quota change, which could contribute to
the observed quota stability. More generally, reviews by @Biais1995 and
@Patterson2007a document many cases where changes in catch
quotas that a management agency set were more modest than changes that
would be recommended just by considering variations in stock abundance.

We use a classic fisheries management question to examine how accounting
for costs of policy adjustment can change optimal policies (see also
@Ludwig1980, @Feichtinger1994, @Wirl1999). We focus on how
harvest quotas for a stochastically varying fish population can be
chosen to maximize the net present value of a fishery. Our formulation
and solution method largely follow Reed's classic treatment on this
question, a treatment repeated widely in bioeconomic textbooks. We 
note that while later work has extended this treatment to deal with 
a variety of other issues (e.g. @Sethi2005; @Singh2006; @McGough2009),
we start from the classical model for simplicity of presentation and analysis.

With his formulation, Reed showed that a constant escapement policy could be
optimal under certain conditions. Such a policy involves choosing annual
quotas that are perfectly responsive to recruitment variation in a fish
stock. In poor recruitment years, the quota is set to zero and no fishing
is allowed. But any time there is a good recruitment pulse, a quota is
set that allows the fishery to exactly compensate through harvesting,
thereby maintaining the optimal escapement level. However, in that
analysis, Reed did not account for any costs of policy adjustment, which
for such a responsive management strategy potentially could be large.

<!-- NOTE perhaps comment on the deviation from constant escapement !-->

As the example in Figure 1 makes clear, management policies will rarely be
as responsive as this constant escapement policy assumes. In this paper,
we explore conditions under which this lack of responsiveness may be a
rationale response on the part of managers. Specifically, we consider a
case where managers seek to balance the benefits in terms of increased
profits from fishing from more finely tracking recruitment variations
with the growing costs associated with adjusting policies quickly to do
so. The policy adjustment costs involved could reflect pure administrative
transaction costs or preferences held by fishermen, fish processing plants
or other stakeholders, for less variable quotas. In seeking to account
for these policy adjustment costs, we recognize that we do not know just
what functional form they should take and that it will require a lot of
empirical work to estimate that. Therefore, we scope three candidate
functional forms that represent different assumptions about how these
costs operate to determine whether the results we obtain are sensitive
to such differences.

Methods
=======

@@

We will assume Beverton-Holt dynamics with multiplicative environmental noise

@@

where $N_t$ the stock size, $h_t$ the harvested level, 
$Z_t$ gives the stochastic shocks, which we
assume are lognormally distributed and $A$ and $B$ are positive
constants.

We assume managers set an annual quota for harvesting $h_t$ (the control
variable) after observing the stock size that year $N_t$ (the state
variable), but while still being uncertain about future environmental
conditions and stock sizes. Through time, this gives a time path of
management actions ${\bf h}=(h_1,h_2,\dots)$ that depend on the stock
sizes that were observed (a state dependent control rule).

We assume that managers choose annual quotas to maximize the expected
net present value (NPV) of the fishery. We take as a base case the
situation where there are no costs associated with policy adjustment and
the managers' objective is

\begin{equation}
\max_{{\bf h}}\mathbf{ E} ( NPV_{0} )=\max_{{\bf h}} \sum_0^\infty \mathbf{E} \left( \displaystyle \frac{\Pi_0(N_t,h_t)}{(1+\delta)^{t+1}} \right) \label{eq:objective}
\end{equation}

where $\mathbf{E}$ is the expectation operator, $\delta>0$ is the
discount rate and $\Pi_0$ is the net revenue from operating the fishery
in a given year. In this base case, we assume the annual net revenue
from fishing is

\begin{equation}
\Pi_0(N_t,h_t) = p h_t -  c_0 E_t \label{eq:annual_netrevenue}
\end{equation}

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t = q E_t N_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. \eqref{eq:annual_netrevenue}, $p$ is the price per unit harvest and
$c_0$ the cost of fishing and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$. (Examples
shown here use $p$= `r price` and $c_0$ = `r c0`, complete code is
provided in the supplementary materials to re-run these analyses with
arbitrarily specified values and replot the figures shown).

Taken together this objective function \eqref{eq:objective} and the state
equation \eqref{eq:state_equation} define a stochastic dynamic programming
problem that we solve using backwards recursion via Bellman's equation.
We denote the resulting state-dependent, optimal control as $h_0^{\ast}$.

We solve this problem on a finite time horizon of $T =$ `r OptTime` using
value iteration [@Mangel1988; @Clark2000].

@@

We compare this base case to three alternative problem formulations,
each reflecting different plausible functional forms that costs of
policy adjustment could take. In each, we assume managers can adjust the
quota set in the fishery $h_t$ in a given year and that any policy
adjustment costs are associated with changes to this control variable.

First we assume that policy adjustment costs are directly proportional
to the magnitude of the change in policy being proposed, such that
larger changes to annual harvesting quotas incur greater policy
adjustment penalties. Specifically, we replace $\Pi_0$ in the $NPV_0$
equation with

@@

Next we continue to assume that policy adjustment costs depend on the
magnitude of the change in policy being proposed. However, we consider a
case where this dependence is nonlinear with big changes in policy being
disproportionately expensive:

@@

Finally, we assume instead that there is a fixed cost associated with
any adjustment of policy, regardless of how large that adjustment might
be

@@

where indicator function $\mathbf{I}(h_t, h_{t-1})$ takes the value 1 if
$h_t=h_{t-1}$ and zero otherwise. For each $\Pi_i$, we can then define a
new objective function $NPV_i$ similar to that in Eqn. \eqref{eq:objective}.

We note that the fixed cost has conceptual analogues to the set-up cost of
@Reed1974 and @Spulber1982 but here the fee is associated with a change
in harvesting ($h_t\neq h_{t-1}$) and not with harvesting *per se*
($h_t>0$) as in those studies.

Taken together with the state equation Eqn. \eqref{eq:state_equation}, each
of these new objective functions defines a different stochastic dynamic
programming problem. Again, we solve them numerically using backwards
recursion. To include costs of policy adjustment, we expand the state
space to include both the current stock size and the management action
taken on the previous time step $(N_t,h_{t-1})$. For each new objective
function $\max NPV_i$, we denote the corresponding optimal control
policy $h_i^*$.

In addition in the Supplementary Material, we compare our results with 
more conventional fisheries economic formulation in which additional
costs are applied to the control variables themselves, as opposed to
adjustments to the controls: $\Pi_4 (N_t,h_t) = p h_t -  c_0 E_t-c_4E_t^2$. 
Additional costs of this form tend to have a smoothing effect on optimal quotas, but also change
the long-term average stock size or quota size that is optimal.

@@

<!--NOTE ci notation is confusing because it does not usually refer to c0 or c4 -->

Each of the policy adjustment cost function is characterized in terms of
a cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect of
each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal net
present value of fishery, broadly following the example of @Bovenberg2008.

<!--NOTE Jake to confirm these are still relevant and needed. 
        We are using a less sophisticated normalization than was 
        once discussed and may or may not need the citations now. -->

Figure 2 shows the calibration graphically. Each curve plots the change
in maximum expected NPV for a given penalty function as the penalty cost
parameter is increased. In each case, the figure shows maximum expected
NPV with policy adjustment costs as a proportion of the maximum expected
NPV available in the basic problem $NPV_0({\bf h_0^*})$ without policy
adjustment costs.

```{r Figure_2, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost, $\\tfrac{NPV_0 - NPV_i}{NPV_0}$.  The horizontal line indicates a value of the stock that is reduced by `r reduction * 100`% from the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=4, fig.height=3, cache = fig.cache}

@@

write.csv(dt, "components/data/100reps.csv")
write.csv(fees, "components/data/figure2.csv")

@@

To compare the impact of penalty functions on optimal management across
the different functional forms, we select penalty cost coefficients that
induce the same reduction in maximum expected NPV. For example, the
dashed vertical lines in Figure 1 map the needed cost coefficient $c_i$,
for each penalty function, such that the fishery is worth `r 100*(1-reduction)`% of its
unconstrained value when optimally managed.

We focus our examination of the models on how the type and severity of
policy adjustment costs affects the optimal management strategy. To
simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the intrinsic rate of increase of
the fish population is 1 and the equilibrium biomass for the equivalent
deterministic model without harvesting is 10; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn. (1)). To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with 
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.

<!-- is chosen such that the
optimal stock size in the analogous deterministic problem is 5% larger
than that corresponding to the maximum sustainable yield in biomass
($c_0=XXX$). -->

Results
=======

@@

```{r Figure_3, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=3, fig.width=5, cache = fig.cache}

@@

ggplot(harvest_fig3, aes(time, value, col=baseline)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Stock & Harvest Dynamics")

write.csv(harvest_fig3, "components/data/figure3.csv")
```

Figure 3 illustrates how the different forms of adjustment cost can impact
the optimal harvest policy. Corresponding stock sizes, which are measured
here before harvest ($h_t$ being determined after observing stock size
$x_t$) and thus show less influence of the policy cost can be seen in the
supplement.  Each panel is generated against the same sequence of random
shocks so that they can be compared directly.  The harvest policy chosen
shows a systematic deviations from the penalty free optimum, depending on
the nature of the cost.  In each case, the optimal solution without any
adjustment cost is shown in against the policy induced by optimization
under the given adjustment penalty (equivalent to a `r 100*reduction`%
reduction in maximum expected NPV) is overlaid in blue.

<!-- NOTE FIXME make this consistent! -->

When considering the first formulation of policy adjustment costs
(linear costs, $\Pi_1$), the first panel of Figure 3 shows a typical pattern, in which
the optimal policy tends to avoid small policy adjustments, resulting in
periods of constant policy followed by sudden bursts of adjustment.
This results in a relatively step-like policy pattern. In contrast, the
second formulation (quadratic costs, $\Pi_2$) disproportionately penalizes large policy
adjustments. The corresponding optimal policy tracks all of the changes
made by the cost-free policy, but with smaller magnitude. This results
in a smoother $h_t$ curve, one that undershoots the larger oscillations
seen in the cost-free optimum in favor of a policy that changes
incrementally each year. Finally, the optimal policy for the third,
'fixed fee' formulation ($\Pi_3$) only makes large adjustments, as one
might expect, because the magnitude of the adjustment made is not
reflected in the resulting penalty.

The comparisons shown in Figure 3 are for one realization (i.e. a
particular sequence of random number draws representing environmental
variability) and are made for a particular magnitude of policy
adjustment costs equal to `r reduction`% of the maximum expected net present value 
in the base case ($NPV_0({\bf h_0^*})$). Figure 4 generalizes to show
how optimal harvest levels and corresponding optimal stock sizes are
affected by the magnitude of the costs of policy adjustment. This time
we show summary statistics across 50 replicate sets of environmental
conditions. We show the impact of policy adjustment costs on the
variance of both harvests and stock sizes through time; on the
autocorrelation of harvests and stock sizes through time. <!-- NOTE Are we showing cross-correlations or not? --> <!-- See figure 4, probably not-->

For example, when small policy adjustment costs cost little but large
adjustments are expensive ($\Pi_2$), we see the smoothing signal that we
might have expected (see also @Ludwig1980 for this particular case). As
cost penalties increase in severity (i.e., $c_2$ increases), the
variance in quotas through time decreases and the autocorrelation of
quotas through time increases. This smoothing of the harvest quotas
through time has knock-on effects for the corresponding optimal stock
sizes, which also become more autocorrelated as the scaling on policy
adjustment costs ($c_2$) is increased. The autocorrelation in stock
sizes here reflects the lag involved in a more modest increase in
harvests taking longer to eliminate the signal of strong recruitment
events, when compared to a more immediately responsive constant
escapement policy.

Interestingly, as suggested by the realization in Fig. 3, including a
fixed cost of policy adjustment increases the variation in harvests
through time. This is the opposite of a smoothing effect. Follow-on
consequences for the other statistics are not as clear in this case.

Finally, the case where policy adjustment costs scale linearly with the
size of the adjustment ($\Pi_1$ in Eqn. 4) appear to be something of a
middle of the road strategy, in that increasing the severity of policy
adjustment costs (increasing $c_1$ has little effect on the variance or
autocorrelation of optimal harvest rates or stock sizes. To reveal the
particular impact of policy adjustment costs of this type requires a
more targeted summary statistic. Specifically, for each run we
calculated the frequency with which the optimal policy involved
maintaining a positive quota across multiple time steps unaltered. This
type of policy is arguably the most commonly observed behavior in TAC
management, but is one that is almost never observed to be part of the
optimal management strategy in the basic model without policy adjustment
costs (Eqn. 3) or when optimizing against $\Pi_2$ or $\Pi_3$.

```{r}
fraction_no_shift <- function(x){

noshift <- x[2:length(x)] - x[1:length(x)-1] == 0
strictly_positive <- (x > 0.01 )[-1] # drop first point, 
who <- noshift & strictly_positive # True whenever consecutive pairs did not shift and the second in the pair was nonzero harvest
 sum(who)/length(who)
}

shifts <- dt[, fraction_no_shift(harvest), by=c("penalty_fn", "replicate")]
ave <- shifts[, mean(V1), by=penalty_fn]

no_shifts <- ave$V1
names(no_shifts) <- ave$penalty_fn
```

While still not common for the particular parameter combinations we examine,
we find that positive unaltered quotas through time are much more likely
to occur when optimizing against $\Pi_1$ where policy adjustment costs
scale linearly with the size of proposed policy changes: in the 100 replicates
of data presented above the harvest policy is strictly positive and identical in consecutive intervals
only `r no_shifts[["L2"]] * 100`% of the time for quadratic costs $\Pi_2$, compared with
`r no_shifts[["L1"]] * 100` with linear costs $\Pi_1$ and `r no_shifts[["fixed"]]*100`%
for fixed costs $\Pi_3$. Moreover, these occurrences increase in frequency as the severity of
these costs ($c_1$) increases.

<!-- to include cross.correlation, drop subset call and facet_wrap by variable. But shows no pattern-->

```{r Figure_4, fig.cap="Variance and autocorrelation of stock dynamics and harvest dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs.", fig.height=8, fig.width=8, cache=fig.cache}
ggplot(subset(figure4_df, statistic != "cross.correlation"), 
       aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ timeseries + statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  xlab("Penalty size (as fraction of NPV0)")

@@

```{r Figure_5, fig.cap="Distribution of profits from fishing and costs paid to adjustment.  Distributions are shown when managing a system in which the adjustment costs present are determined by the model indicated, and compared to the same costs and profits of managing the same system using a Reed model, which ignores adjustment costs, and thus tends to pay much higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).", fig.width=6, fig.height=5, dependson="histogram-calc"}

@@

Next we examine the consequences either of ignoring policy adjustment
costs when they are present or assuming they are present when they are
not. To do so, it helps to break-down the different contributions to the
maximum expected $NPV_i$ when employing the optimal policy $\bf{h_i}^*$
versus when employing the policy that would be optimal were there no
costs to policy adjustment $\bf{h_0}^*$. For example, if we take the
first of the adjustment penalty forms, $\Pi_1$

@@

Histograms for the different contributions labelled $[1]-[4]$ here when
taken across 100 replicate runs are shown are shown in Figure 5.
<!--(contribution $[1]$ red, $[2]$ green, $[3]$ blue, and $[4]$ mauve).--> The
cost of assuming policy adjustments are present when in fact they are
absent is given by subtracting the contribution to the first infinite
sum from term $[1]$ from the contribution to the second infinite sum of
term $[3]$. (Note that convergence properties
required to do this are met because we assume a positive discount rate).
This cost is positive because the optimal controls recommended when
assuming policy adjustment costs are present do not track variations in
stock size as closely. Therefore, this harvesting strategy provides
lower net revenue dockside.

In contrast, the cost of ignoring policy adjustment costs when in fact
they are present is given by subtracting the second infinite sum from
the first (contributions $([1]-[2])-([3]-[4]))$ in the equations. These
costs are positive because the lost revenues from not tracking stock
variations as closely are smaller than the penalties of policy
adjustment that are incurred if managers try to track stock variations
more tightly.

```{r Figure_6, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs."}

who <- c("penalty_fn", "ignore_fraction", "assume_fraction", "reduction")
table1 <- arrange(error_df[who], reduction) 
names(table1) = c("penalty.fn", "ignoring", "assuming", "reduction")
table1_long <- melt(table1, id = c('penalty.fn', 'reduction'))
table1_long$reduction <- as.factor(table1_long$reduction)
table1_long <- subset(table1_long, reduction != "0.3")
ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
write.csv(table1_long, "components/data/figure6.csv")
```

Figure 6 compares these different contributions, namely the $NPV$ lost by
assuming policy adjustment costs when none are present (first entry in
each cell, $[3]-[1]$ in the equations)) and the $NPV$ lost when
ignoring policy adjustment costs when they are present (second entry in
each cell of the table, contributions $([1]-[2])-([3]-[4]))$ in the
equations) across the three different functional forms. Specifically,
the table shows how these values change as we vary the severity of
policy adjustment costs. In each case, values are shown as percentages
of the maximum expected $NPV$ in the base case where no policy
adjustment costs apply or are assumed to apply $NPV_0({\bf h_0^*})$.

We find a pronounced asymmetry when comparing the impact on the net
present value of the fishery of assuming policy adjustment costs are
present when they are absent with the impact of assuming they are absent
when they are present. For all three types of policy adjustment cost,
the impact on NPV of ignoring these costs if they are in fact present is
much bigger than that incurred by fallaciously managing as if they are
present when they do not occur. This asymmetry arises, because there is
less difference in the dockside revenue under the two management
recommendations than there is in the potential policy adjustment costs
themselves, as is evident in for the particular parameters shown in
Figure 5.

We also see that the impacts on net present value of the fishery of
incorrectly assuming or ignoring policy adjustment costs are most severe
when these costs are assumed to scale quadratically with the size of the
policy change being implemented ($\Pi_2$).

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure 7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10%
less than optimal free. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

For fixed costs ($\Pi_3$), the reverse is true.  Because fixed costs increase the
volatility of the optimal solution while $\Pi_2$ costs decrease it, simply
ignoring the cost of adjustment all together results in a higher value
than either assuming costs are quadratic ($\Pi_2$) when they are fixed ($\Pi_3$),
or assuming they are fixed when they are quadratic.

Nevertheless, assuming quadratic costs is always the most conservative
option -- scenarios in which true costs are quadratic (regardless of what
cost structure is assumed by the policy) do significantly worse than
those that are not (i.e. all cases that assume quadratic but actually
have linear or fixed costs).

These patterns hold across the different size reductions shown (from 10%
to 25%), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

```{r Figure_7, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.  Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel should lose only 10% of its net present value to adjustment costs if it had the correct form of those costs."}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
write.csv(table2_long, "components/data/figure7.csv")
```

Discussion
==========

<!-- Pasted from external word doc, see Boettiger_Discussion_draft.docx for latest version of discussion -->

Policy-makers managing ecological systems that vary in space and time
must evaluate how much of that variation to reflect in management
recommendations. Fine-tuning a policy to respond to frequent variations
in ecological dynamics may incur increased transaction costs associated
with constantly revisiting past policy decisions. A more pragmatic
approach would be one that balances benefits from responding to frequent
variations in ecological conditions with the increased transaction costs
involved. As a first step towards exploring these ideas, we revisited
a classic problem from bioeconomics concerning the optimal management
of a fish stock subject to stochastically varying recruitment [@Reed1979; @Clark2010]. 
We examined how optimal policy recommendations, here
annual catch quotas, changed when accounting for costs associated with
policy adjustment and what the implications of following these policy
recommendations would be for the exploited population. We also compared
how the value of the fishery is affected by managers either under-
or over-estimating the importance of policy adjustment costs of this type.

Estimating policy adjustment costs and how they respond to the size of
proposed policy changes would be empirically challenging. Recognizing
this fact, in our analyses, we compared different plausible forms that
these costs might take. We compared two cost structures where we assumed
the magnitude of policy adjustment costs increased with the magnitude
of the change in policy being proposed with an alternative formulation
in which we assumed there was a fixed cost associated with making any
change to current policy. While each formulation clearly provides only a
very phenomenological representation of the policy setting environment,
we believe that each enables us to explore meaningful differences in how
policy adjustment costs might operate. That being said, in the real-world
we might expect the different types of adjustment cost to operate in
combination. Interestingly, there are analogues between how we have
represented different types of policy adjustment costs in our models
and regularization techniques that are sometimes used in model fitting
to address concerns about over-specifying models (e.g. ridge or lasso
regression or total variation denoising approaches
[@Rudin1992]).

<!-- NOTE

Have wondered about cutting this, but does broaden things out a bit from
just econ referencing and context.

Also more people will have heard about ridge and lasso, but I actually
think they're not as relevant as the total variation suggestion, because
it applies to the gradient whereas they do not. So really a more accurate
statement is that the difference between lasso and ridge and the total
variation approach is in some ways analogous to the difference between
costs on h or E and costs on dh/dt or dE/dt. But that's way too many
words now

Carl B - when it comes to references, you may have others you would
prefer here

-->

>>>>

---
title: "Fisheries management when quotas are costly to change"
date: "07/24/2014"
output: ioslides_presentation

author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviera
    affiliation: UTK
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTK

@@

---

@@

library("methods")
library("knitr")
basename <- "presentation"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

@@

```{r Figure_1, fig.cap="Quota on Blue Fin Tuna Catch", dev=c("svg", "png")}
tuna <- read.csv("components/data/tuna.csv")
ggplot(tuna, aes(year, quota)) + geom_point()
```

@@

@@

Optimization
------------

$$\max_{{\bf h}}\mathbf{ E} ( NPV_{0} )=\max_{{\bf h}} \sum_0^\infty \mathbf{E} \left( \displaystyle \frac{\Pi_0(N_t,h_t)}{(1+\delta)^{t+1}} \right)$$

$$\Pi_0(N_t,h_t) = p h_t -  c_0 E_t $$

@@

We replace $\Pi_0$ in the $NPV_0$ equation with

@@

@@

@@

@@

```{r Figure_2, fig.cap="",  dev=c("svg", "png")}

@@

@@

@@

```{r Figure_3, dependson="tidy", fig.cap="", dev=c("svg", "png")}

@@

ggplot(harvest_fig3, aes(time, value, col=baseline)) +
  geom_line(lwd=1) +
  facet_grid(penalty_fn~., labeller = labeller) + 
  labs(x="time", y="stock size", title = "Example Stock & Harvest Dynamics")  
```

## Figure 4

```{r Figure_4, fig.cap="", dev=c("svg", "png")}
ggplot(subset(figure4_df, statistic != "cross.correlation"), 
       aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ timeseries + statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  xlab("Penalty size (as fraction of NPV0)")

@@

```{r Figure_5, fig.cap="",  dependson="histogram-calc", dev=c("svg", "png")}

@@

## Penalty

@@

## Figure 6

```{r Figure_6, fig.cap="", dev=c("svg", "png")}

who <- c("penalty_fn", "ignore_fraction", "assume_fraction", "reduction")
table1 <- arrange(error_df[who], reduction) 
names(table1) = c("penalty.fn", "ignoring", "assuming", "reduction")
table1_long <- melt(table1, id = c('penalty.fn', 'reduction'))
table1_long$reduction <- as.factor(table1_long$reduction)
table1_long <- subset(table1_long, reduction != "0.3")
ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
```

## Mismatches

```{r Figure_7, fig.cap="", dev=c("svg", "png")}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
```
